### 2014-03-17 paper/designspace.tex
bjoern's comments

\section{The Design Space of Tubes}

@@

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/tubespace.png}
\caption{The design space of tubes.  Tube types, media, topologies, and design are discussed more fully in the text.  \tovi{May want star, tree added to 
your figure
}
}
\label{fig:tubespace}
\end{figure}

\subsection{Types of Tubes}\tovi{Maybe rename types to 
"openings"?}

Tubes can come in four types: open, return, semi-closed, and fully enclosed.  Each of these types offers distinct interaction capabilities (see Figure \ref{fig:tubespace}).  Definitionally, in the following paragraphs we refer to a tube which is ``open on the user side'' to have its contents physically accessible to an interacting user at interaction time.  This includes tubes which function as fountains, but it does not include tubes whose ends have been capped to keep water inside (whether this capping was done at print time, or later as a manual modification).  Tubes whose contents are not physically accessible to the user are ``closed on the user side''.  A tube which is ``open on the system side'' can have its contents manipulated by a machine, e.g., by resonating, pressurizing, or passing electricity through those contents.  The system does not need to have physical access to the contents in order to manipulate them.

@@

Semi-closed tubes are open at the system end (for control of the enclosed medium) and closed at the user end.  The closed interface is interesting when it fabricated from a mobile material: for example, a series of tubes terminated in thin rubber membranes on the user side can be actuated by an air pump to create haptic feedback, or a hinged cap which can be blown open by air pressure.  Without a printer capable of fabricating flexible material, a maker could affix a balloon to an open tube's end to behave similarly; another possibility is to use semi-closed tubes as audio-generating resonance chambers.

A fully enclosed tube has no openings on either the system side or the user side.  Fully enclosed tubes can be used as resonance chambers (e.g., for object identification), or as air bubbles (e.g., as used in \cite{Willis-printedoptics} for internal display). \tovi{used as a container for water 
or other particles that would 
otherwise fall out?}  \valkyrie{I re-worded above.  please check.}  Their physical design space is very limited, as any support material required to create their internal geometry cannot be accessed for removal.

@@

``Gas'' comprises all compressible fluids.  Use of fluid pressure inside tubes can create haptic feedback at semi-closed interfaces, or the gases can be used as carriers for scents or fog.  As in \cite{Slyper-pressure}, structures can be engineered to change in air pressure when manipulated correctly (e.g., a spiral that changes pressure when twisted, but not pressed), and thus fluid pressure can also be used as an input. \tovi{I realize "fluid" pressure is 
technically correct, but might 
confuse readers, since this is 
the gas pararaph.}

@@

\tovi{Again, this raises the question of how we are applying the design 
space - is it based on its properties when it is printed, or based on 
any post-print modifications}\valkyrie{re-worded above.  I think we are basing it on status at ``interaction time"}Tubes need not have hollow centers: in the case where routed tubes are filled with solid material---in particular, a solid material different from the model material---, interactions such as those in \cite{Willis-printedoptics} are possible.

@@

\subsection{Tube Topologies}

Tube network topologies enable different types of interactions.

Splitting or mixing tubes offer flexibility in output.  If a maker wished to create a painting device, she might wish to have two system-side tubes feeding in primary-colored red and blue paints which mix in varying ratios, allowing their pigments to combine before purple exits from the device (see Figure \ref{fig:tubespace}).  Splitting can also be useful, for example if our maker wants red paint output in two locations from her painting device, she could have one system-side tube, but split the tube into two (see Figure \ref{fig:tubespace}).\tovi{How about multiple in/
multiple out (in contrast to 
branching/splitting which is 
single in multiple out)
} \tovi{This seciton should refer to the 
possibilities, not to what we 
actually did.}

Star and tree topologies are extensions of the splitting and mixing primitives.  Using a star topology in which the tubes were filled with conductive paint, we created toys with several touch-sensitive areas, see Figure \ref{fig:toys}.

\subsection{Design of Tubes}

Tube designs may emphasize either their exterior features (connection points) or their interior paths.  These two features lead to different kinds of interfaces.  An example interface that focuses on the exterior connection points is the touch sensitive toy in Figure \ref{fig:toy}, where tubes must exit the toy at the eyes, ears, tail, etc.  An interface focused on the internal path of the tube is the neon sign in Figure \ref{fig:neon}: output is based on the shape of the tube.

@@

>>>>

\section{The Design Space of Tubes}
\bjoern{We should be careful with our terminology - apparently pipes and tubes have different connotations. Tubes are measured by outer diameter (OD) and tend to be structural, while pipes are measured by inside diameter (ID) and tend to be used to transport media inside them. At the least, we should mention that different communities have these different connotations. I think that in our case, we're thinking mostly of the 3D print as the structural object that we're then tunneling through with {\em pipes}.}

@@

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/tubespace.png}
\caption{The design space of tubes.  Tube types, media, topologies, and design are discussed more fully in the text.}
\label{fig:tubespace}
\end{figure}

\subsection{Types of Tubes}

Tubes can come in four types: open, return, semi-closed, and fully enclosed.  Each of these types offers distinct interaction capabilities (see Figure \ref{fig:tubespace}).

@@

Semi-closed tubes are open at the system end (for control of the enclosed medium) and closed at the user end.  We believe this closed interface is most interesting when it is fabricated from a mobile material: for example, a series of tubes terminated in thin rubber membranes on the user side can be actuated by an air pump to create haptic feedback.  Without a printer capable of fabricating flexible material, a maker could affix a balloon to an open tube's end to behave similarly; another possibility is to use semi-closed tubes as audio-generating resonance chambers.

While possible to create, a tube which is open on the user side and closed on the system side is outside our focus on computer-mediated interaction, and we therefore do not discuss these tubes.

A fully enclosed tube has no openings on either the system side or the user side.  Fully enclosed tubes can be used as resonance chambers (e.g., for object identification), or as air bubbles (e.g., as used in \cite{Willis-printedoptics} for internal display).  Their physical design space is very limited, as any support material required to create their internal geometry cannot be accessed for removal.

@@

``Gas'' comprises all compressible fluids.  Use of fluid pressure inside tubes can create haptic feedback at semi-closed interfaces, or the gases can be used as carriers for scents or fog.  As in \cite{Slyper-pressure}, structures can be engineered to change in air pressure when manipulated correctly (e.g., a spiral that changes pressure when twisted, but not pressed), and thus fluid pressure can also be used as an input.

@@

Tubes need not have hollow centers: in the case where routed tubes are filled with solid material---in particular, a solid material different from the model material---, interactions such as those in \cite{Willis-printedoptics} are possible.

@@

\subsection{Topological variations}

Tube topology enables different types of interactions.

Splitting or mixing tubes offer flexibility in output.  If a maker wished to create a painting device, she might wish to have two system-side tubes feeding in primary-colored red and blue paints which mix in varying ratios, allowing their pigments to combine before purple exits from the device (see Figure \ref{fig:tubespace}).  Splitting can also be useful, for example if our maker wants red paint output in two locations from her painting device, she could have one system-side tube, but split the tube into two (see Figure \ref{fig:tubespace}).

Star and tree topologies are extensions of the splitting and mixing primitives.  Using a star topology in which the tubes were filled with conductive paint, we created a toy with several touch-sensitive areas, see Figure \ref{fig:toy}.

\subsection{Features of Tubes}

Tubes may emphasize either their exterior features (connection points) or their interior paths.  These two features lead to different kinds of interfaces.  An example interface that focuses on the exterior connection points is the touch sensitive toy in Figure \ref{fig:toy}, where tubes must exit the toy at the eyes, ears, tail, etc.  An interface focused on the internal path of the tube is the neon sign in Figure \ref{fig:neon}: output is based on the shape of the tube.

@@

\valkyrie{this section includes related work.  is that ok?  it's in the table, and I think we should cover the related input techniques here rather than in the designated related work section (which I think is more suited to fabrication and routing stuff).}

### 2014-03-17 paper/introduction.tex
bjoern's comments

\section{Introduction}
Makers, as well as professional designers, leverage 3D printers as tools for design work.  A wide array of objects, ranging from bicycle helmets to jewelry to video game controllers, are now prototyped or even manufactured using these machines.  However, most devices fabricated by 3D printers are passive: accessible printers are not yet capable of creating integrated active systems.

Willis, et al., ``envision a future world where interactive devices can be printed rather than assembled; a world where a device with active components is created as a single object, rather than a case enclosing circuit boards and individually assembled parts" \cite{Willis-printedoptics}.  This is a vision we eagerly share: 3D printers are capable of creating arbitrary geometries not feasible to manufacture using traditional processes, and we see these capabilities being underutilized by makers and designers.  We see many opportunities to increase the interactivity of 3D prints using today's printer technology.
\tovi{Might want some examples of what possibilities this opens up}

Using specially modified printers or extra machinery, it is possible to create electronics on the surface of 3D prints \cite{optomec} \cite{Sells-reprap}.  These techniques, however, require a high capital investment and technical expertise.  In addition, they lack flexibility: the printed circuits must be routed in 2D to be created on the objects' manifold exteriors, and they can only be used to create electronic circuits. \valkyrie{what?}  \tovi{I dont follow this paragraph - I think you want  to briefly summarize what has been previousl done in interactive 3d prints, and hihglight the shortcommings and remianing challenges}

@@

In this paper, we propose a new technique, where interior channels, such as hollow champers or solid, are integrated into 3D models. Such channels introduces a entire design space of opportunities for adding interactvity to 3D printed objects. For example, copper material can fill the channels, to allow for standard electronic compoents to be easily integrated after the printout; air can be pumped through to create tactile output; or, Electroluminescent wire can be passed through to create interactive visual output.

Our work makes the following contributions to the field of interactive digitial fabircation. First, ... deisgn space...Second ... modeling tool and algorithms ... Third showcase a set of examples, enabled by our modeling tool.

>>>>

\section{Introduction}
Makers, as well as professional designers, leverage 3D printers as tools for design work.  A wide array of objects, ranging from bicycle helmets to jewelry to video game controllers, are now prototyped or even manufactured using these machines.  However, most devices fabricated by 3D printers are passive: accessible printers are not yet capable of creating integrated active systems. \bjoern{not sure I understand phrase after the colon. Rewrite and explain what you mean in more detail?}

Willis, et al., ``envision a future world where interactive devices can be printed rather than assembled; a world where a device with active components is created as a single object, rather than a case enclosing circuit boards and individually assembled parts" \cite{Willis-printedoptics}.  This is a vision we eagerly share: 3D printers are capable of creating arbitrary geometries not feasible to manufacture using traditional processes, and we see these capabilities being underutilized by makers and designers.  We see many opportunities to increase the interactivity of 3D prints using today's printer technology.
\tovi{Might want some examples of what possibilities this opens up}
\bjoern{I think we're doing something a little different from Karl's vision - in all of our examples, you need to hook up extra sensors, actuators, or lights to the 3D-printed object. So there's always still assembly required. We're not really printing interactive objects in a single pass, but we're offering new ways of adding interactivity to 3D printed objects.

I'd argue the principal novelty of our work is to {\em remove material} from a model to form pipes and other cavities. We demonstrate how designers/makers can then use these pipes, e.g. by filling them with various media to enable input, display and tactile feedback. This {\em subtractive} approach is complementary to {\em additive} approaches that try to print different materials such as conductors. While our approach requires manual assembly afterwards, it opens up a large useful design space on commonly available 3D printers.}

Using specially modified printers or extra machinery, it is possible to create electronics on the surface of 3D prints \cite{optomec} \cite{Sells-reprap}.  These techniques, however, require a high capital investment and technical expertise.  In addition, they lack flexibility: the printed circuits must be routed in 2D to be created on the objects' manifold exteriors, and they can only be used to create electronic circuits. \valkyrie{what?}  \tovi{I dont follow this paragraph - I think you want  to briefly summarize what has been previousl done in interactive 3d prints, and hihglight the shortcommings and remianing challenges} \bjoern{agree. Also, probably too much focus on electronics since it's only one of a series of examples we'll demonstrate.}

@@

In this paper, we propose a new technique, where interior channels, such as hollow chambers or solid \bjoern{?}, are integrated into 3D models. Such channels introduces a entire design space of opportunities for adding interactvity to 3D printed objects. For example, copper material can fill the channels, to allow for standard electronic components to be easily integrated after the printout; air can be pumped through to create tactile output; or, Electroluminescent wire can be passed through to create interactive visual output. \bjoern{this is good - I'd make this the central argument of the intro and move it up earlier.}

Our work makes the following contributions to the field of interactive digitial fabrication. First, ... design space...Second ... modeling tool and algorithms ... Third showcase a set of examples, enabled by our modeling tool. \bjoern{I'd say more about the software contribution in the abstract. Maybe that just requires expanding these notes into a full paragraph.}
